---
title: "The amenity basket"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The amenity basket}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- closecity::close_client(api_key = Sys.getenv("CLOSECITY_KEY"))
```

A city planner wants every resident to be able to walk to a basket of six everyday
amenities: a supermarket, a library, a park, a frequent-transit stop, a restaurant,
and a cafe. This tutorial measures how many residents already have that, and shows
where the gaps are. The idea follows
[this analysis](https://nathenry.com/writing/2023-02-07-seattle-walkability.html),
here applied to Richmond, Virginia.

*Running this tutorial uses about 3,500 tokens.*

## Set up

Read the six category ids from the free catalog, and turn the city name into a
centre point and a boundary.

```r
library(closecity)
library(sf)
close <- closecity::close_client(api_key = "ck_live_your_key")   # use your own key here
```

```{r}
amenity_types <- close$destination_types()
ids <- setNames(amenity_types$dest_type_id, amenity_types$label)

basket <- c(
  supermarket = ids[["grocery_stores"]],
  library = ids[["libraries"]],
  park = ids[["parks"]],
  transit = ids[["frequent_transit"]],
  restaurant = ids[["restaurants"]],
  cafe = ids[["cafes"]]
)

city <- close$places(q = "Richmond")[1, ]
city_boundary <- close$place_boundary(geoid = city$geoid)
```

## Pull the blocks, with population

`$blocks_query()` reads every page: the walk time from every block to each of the
six categories, plus each block's population, as one sf table. To keep this
tutorial cheap we take the central blocks within a radius; `$place_blocks(geoid = city$geoid)`
pulls **every** block in the city the same way (at a higher token cost).

```{r}
blocks <- close$blocks_query(
  center = list(lon = city$lon, lat = city$lat), radius_m = 2500,
  mode = "walk", type = unname(basket), include_population = TRUE
)

# One row per block, for population and for mapping.
one_per_block <- blocks[!duplicated(blocks$geoid), ]
total_pop <- sum(one_per_block$population)
```

## Coverage, one amenity at a time

For each amenity, a block counts as covered when it is within a 15-minute walk. Add
up the population of the covered blocks.

```{r}
for (name in names(basket)) {
  covered <- unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                                   blocks$travel_time <= 15])
  pop <- sum(one_per_block$population[one_per_block$geoid %in% covered])
  cat(sprintf("%-11s %3.0f%%\n", name, 100 * pop / total_pop))
}
```

Parks and restaurants tend to be everywhere; supermarkets and libraries are usually
the hardest to reach. Map one amenity — every block shown, the covered ones
highlighted, the city boundary behind.

```{r}
near_library <- unique(blocks$geoid[blocks$dest_type_id == basket["library"] &
                                      blocks$travel_time <= 15])
one_per_block$has_library <- one_per_block$geoid %in% near_library
closecity::close_map(x = one_per_block, highlight = "has_library", color = "#058040",
                     boundary = city_boundary)
```

## The 15-minute-city score

Count, for each block, how many of the six amenities are within a 15-minute walk.
That score, from 0 to 6, is the map planners reach for; blue marks the best-served
blocks. It reuses the data you already pulled, so it costs nothing more.

```{r}
covered <- blocks[blocks$travel_time <= 15, ]
score <- tapply(covered$dest_type_id, covered$geoid, function(x) length(unique(x)))
one_per_block$score <- as.integer(score[one_per_block$geoid])
one_per_block$score[is.na(one_per_block$score)] <- 0L
closecity::close_map(x = one_per_block, fill = "score", boundary = city_boundary)
```

## Who can reach all six

A block is fully covered only if all six amenities are within 15 minutes. Fold the
per-amenity covered sets together.

```{r}
covered_sets <- lapply(names(basket), function(name) {
  unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                        blocks$travel_time <= 15])
})
covered_all <- Reduce(intersect, covered_sets)

basket_pop <- sum(one_per_block$population[one_per_block$geoid %in% covered_all])
cat(sprintf("All six amenities: %.0f%% of residents\n", 100 * basket_pop / total_pop))

one_per_block$full_basket <- one_per_block$geoid %in% covered_all
closecity::close_map(x = one_per_block, highlight = "full_basket", color = "#f36e21",
                     boundary = city_boundary)
```

## Which amenity to add first

Look at the residents who are not yet fully covered, and count how many of them are
missing each amenity. The amenity that the most people lack is the one to add first.

```{r}
uncovered <- setdiff(one_per_block$geoid, covered_all)

for (name in names(basket)) {
  covered <- unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                                   blocks$travel_time <= 15])
  lacking <- setdiff(uncovered, covered)
  pop <- sum(one_per_block$population[one_per_block$geoid %in% lacking])
  cat(sprintf("%-11s %6.0f residents would gain access\n", name, pop))
}
```

## Who is one or two amenities away

The residents worth targeting first are the ones *almost* there — a block that has
five of the six is a much easier win than one that has none. Build a block-by-amenity
coverage grid, count how many amenities each block is missing, and for the blocks
short by just one or two, break down which amenities are the gap.

```{r}
covered_by <- lapply(names(basket), function(name) {
  unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                        blocks$travel_time <= 15])
})
names(covered_by) <- names(basket)

# One logical row per block: is each amenity within 15 minutes?
have <- vapply(covered_by, function(s) one_per_block$geoid %in% s,
               logical(nrow(one_per_block)))
n_missing <- rowSums(!have)
gap <- apply(!have, 1, function(row) paste(names(basket)[row], collapse = " + "))

# Residents who are short by exactly one or two amenities.
almost <- n_missing %in% c(1, 2)
almost_pop <- sum(one_per_block$population[almost])
cat(sprintf("%.0f%% of residents are one or two amenities short of the full basket.\n\n",
            100 * almost_pop / total_pop))

pop_by_gap <- tapply(one_per_block$population[almost], gap[almost], sum)
pop_by_gap <- sort(pop_by_gap, decreasing = TRUE)
for (g in names(pop_by_gap)) {
  cat(sprintf("  missing %-22s %3.0f%%\n", g, 100 * pop_by_gap[g] / almost_pop))
}
```

## Site a new supermarket

The counts above say *which* amenity to add; the next question is *where*. Take a
candidate site on land north of the James River and ask how many residents would
newly gain a supermarket within a 15-minute walk if one opened there. A
`direction = "to"` isochrone gives exactly the blocks that could reach the site on
foot in 15 minutes.

```{r}
site_lon <- -77.437
site_lat <- 37.548
reachable <- close$isochrone(lon = site_lon, lat = site_lat, mode = "walk",
                             direction = "to", minutes = 15, format = "blocks")

near_supermarket <- unique(blocks$geoid[blocks$dest_type_id == basket["supermarket"] &
                                          blocks$travel_time <= 15])
newly_served <- setdiff(reachable$geoid, near_supermarket)
gain_pop <- sum(one_per_block$population[one_per_block$geoid %in% newly_served])
cat(sprintf("A supermarket here would newly serve %.0f residents\n", gain_pop))
```

Map the whole city and highlight the blocks that would newly gain access — the
population this one site reaches.

```{r}
one_per_block$newly_served <- one_per_block$geoid %in% newly_served
closecity::close_map(x = one_per_block, highlight = "newly_served", color = "#e8590c",
                     boundary = city_boundary)
```
