---
title: "The amenity basket"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The amenity basket}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

A city planner wants every resident to be able to walk to a basket of six everyday
amenities: a grocery store, a library, a park, a frequent-transit stop, a restaurant,
and a cafe. This tutorial measures how many residents already have that, and shows
where the gaps are. The idea follows
[this analysis](https://nathenry.com/writing/2023-02-07-seattle-walkability.html),
here applied to Richmond, Virginia.

*Running this tutorial uses about 3,500 tokens.*

## Set up

Read the six category ids from the free catalog, and turn the city name into a
centre point.

```r
library(closecity)
library(sf)
close <- close_client("ck_live_your_key")   # use your own key here
```

```{r}
types <- close$destination_types()
ids <- setNames(types$dest_type_id, types$label)

basket <- c(
  grocery = ids[["grocery_stores"]],
  library = ids[["libraries"]],
  park = ids[["parks"]],
  transit = ids[["frequent_transit"]],
  restaurant = ids[["restaurants"]],
  cafe = ids[["cafes"]]
)

city <- close$places("Richmond")[1, ]
```

## Pull the blocks, with population

`$records()` reads every page of a block query: the walk time from every block near
downtown to each of the six categories, along with each block's population. The
result is one sf table.

```{r}
blocks <- close$records(
  "blocks_query",
  center = list(lon = city$lon, lat = city$lat), radius_m = 2500,
  mode = "walk", type = unname(basket), include_population = TRUE
)

# One row per block, for population and for mapping.
one_per_block <- blocks[!duplicated(blocks$geoid), ]
total_pop <- sum(one_per_block$population)
```

## Coverage, one amenity at a time

For each amenity, a block counts as covered when it is within a 15-minute walk. Add
up the population of the covered blocks.

```{r}
for (name in names(basket)) {
  covered <- unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                                   blocks$travel_time <= 15])
  pop <- sum(one_per_block$population[one_per_block$geoid %in% covered])
  cat(sprintf("%-11s %3.0f%%\n", name, 100 * pop / total_pop))
}
```

Parks and restaurants tend to be everywhere; groceries and frequent transit are
usually the hardest to reach. Map one amenity to see the pattern.

```{r}
near_transit <- unique(blocks$geoid[blocks$dest_type_id == basket["transit"] &
                                      blocks$travel_time <= 15])
one_per_block$has_transit <- one_per_block$geoid %in% near_transit

plot(one_per_block["has_transit"], pal = c("#eef0f7", "#058040"), border = NA)
```

## The 15-minute-city score

Count, for each block, how many of the six amenities are within a 15-minute walk.
That score, from 0 to 6, is the map planners reach for. It reuses the data you
already pulled, so it costs nothing more.

```{r}
covered <- blocks[blocks$travel_time <= 15, ]
score <- tapply(covered$dest_type_id, covered$geoid, function(x) length(unique(x)))
one_per_block$score <- as.integer(score[one_per_block$geoid])
one_per_block$score[is.na(one_per_block$score)] <- 0L

plot(one_per_block["score"], pal = hcl.colors(7, "viridis"), border = NA)
```

## Who can reach all six

A block is fully covered only if all six amenities are within 15 minutes. Start with
every block and remove the ones that miss any amenity.

```{r}
covered_all <- one_per_block$geoid
for (name in names(basket)) {
  covered <- unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                                   blocks$travel_time <= 15])
  covered_all <- intersect(covered_all, covered)
}

basket_pop <- sum(one_per_block$population[one_per_block$geoid %in% covered_all])
cat(sprintf("All six amenities: %.0f%% of residents\n", 100 * basket_pop / total_pop))

one_per_block$full_basket <- one_per_block$geoid %in% covered_all
plot(one_per_block["full_basket"], pal = c("#eef0f7", "#f36e21"), border = NA)
```

## Which amenity to add first

Look at the residents who are not yet fully covered, and count how many of them are
missing each amenity. The amenity that the most people lack is the one to add first.

```{r}
uncovered <- setdiff(one_per_block$geoid, covered_all)

for (name in names(basket)) {
  covered <- unique(blocks$geoid[blocks$dest_type_id == basket[name] &
                                   blocks$travel_time <= 15])
  lacking <- setdiff(uncovered, covered)
  pop <- sum(one_per_block$population[one_per_block$geoid %in% lacking])
  cat(sprintf("%-11s %6.0f residents would gain access\n", name, pop))
}
```

Map the uncovered blocks to see where new amenities would do the most good.

```{r}
one_per_block$uncovered <- one_per_block$geoid %in% uncovered
plot(one_per_block["uncovered"], pal = c("#eef0f7", "#202a5b"), border = NA)
```
