---
title: "Get started with closecity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with closecity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Examples run when a key is present (on the docs site), and are shown but not run
# otherwise (on CRAN, and locally without a key). The displayed client uses a
# placeholder; the real one is built here from an environment variable.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

`closecity` reads the Close API: travel times from every US census block to nearby
places, on foot, by bike, and by public transit. This vignette is a short tour.
The three tutorials go further.

## Words you will see

A few terms come up throughout:

- **Census block.** The smallest area the Census Bureau publishes. Each one has a
  15-digit id, its **GEOID**.
- **Destination type.** A category of place, such as grocery stores or libraries.
  Every type has a numeric id.
- **Mode.** How someone travels: walk, bike, or transit.
- **Isochrone.** The area reachable from a point within a time limit, as a polygon.
- **Catchment.** The reverse: every block that can reach a given place.

## Build a client

You make every request through a client object.

```r
library(closecity)
close <- close_client("ck_live_your_key")   # use your own key here
```

The catalog and lookup routes are free, so `close_client()` with no key also works
for those.

```{r}
close$modes()
```

## Look things up instead of guessing

Two free calls save you from memorising codes. Both come back as data frames, so
you filter and index them the usual way: read the numeric id for a category from
the catalog, and turn a city name into a GEOID and a centre point.

```{r}
types <- close$destination_types()
grocery_id <- types$dest_type_id[types$label == "grocery_stores"]

providence <- close$places("Providence")[1, ]
providence$geoid
```

## Make a call and map it

Routes with geometry return an [sf](https://r-spatial.github.io/sf/) object, so you
can map the result straight away.

```{r}
groceries <- close$pois_search(lat = providence$lat, lon = providence$lon,
                               radius_m = 1500, type = grocery_id)
plot(st_geometry(groceries), pch = 19, col = "#202a5b")
```

## Choose an output

Every route returns tabular data by default. The `output` setting controls the
shape: `"spatial"` (the default) gives an sf object where geometry applies and a
data frame otherwise; `"tabular"` gives a data frame everywhere and never downloads
block boundaries; `"raw"` gives the underlying reply. Set it on the client, or pass
`output =` to a single call.

```{r}
close$output <- "raw"
summary <- close$block_summary("440070008001068", mode = "walk")
summary$results
```

```{r, include = FALSE}
close$output <- "spatial"
```

## Handle errors

Failed requests raise a classed condition. Catch the base `close_api_error`, or a
specific one.

```{r}
tryCatch(
  close$block_summary("000000000000000"),
  close_api_error = function(e) message(sprintf("%s (%d)", e$slug, e$status))
)
```
