---
title: "Get started with closecity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with closecity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`closecity` is a thin client over the [Close API](https://api.close.city) — travel
times from every US census block to nearby points of interest, by walking, biking,
and public transit.

## A client

```{r}
library(closecity)

# The catalog and health routes are free; data routes need a key from
# https://account.close.city
close <- close_client("ck_live_your_key_here")
```

## Free catalog and lookup (no key)

```{r}
close <- close_client()
close_modes(close)$data$modes
close_last_updated(close)$data$last_updated

# Resolve a city name to its census place GEOID + centroid:
close_places(close, "Providence")$results[[1]]

# The destination-type ids the data routes filter on:
types <- close_destination_types(close)$data$destination_types
```

Grocery is `30`, restaurants `27`, cafes `31`, libraries `43`, parks `63`, and
frequent-transit stops `61`.

## One metered call

```{r}
summary <- close_block_summary(close, "250173523004004", mode = "walk")
for (row in summary$results) {
  cat(row$dest_type_id, row$mode, row$travel_time, "\n")
}

# Metering is surfaced on every metered reply:
cat(summary$tokens_charged, "charged;", summary$tokens_remaining, "left\n")
```

Metering is **1 token per returned row** (minimum 1 per request); isochrones are the
exception at **10 tokens per contour**. A `304` revalidation is free.

## Pagination

Paginated endpoints take a `cursor`; loop them with `close_records()`:

```{r}
pois <- close_records(close_pois_search, close,
                      lat = 41.823, lon = -71.412, radius_m = 1500, type = 31)
length(pois)
```

## Conditional requests (free revalidation)

```{r}
first <- close_block_summary(close, "250173523004004")
again <- close_block_summary(close, "250173523004004", if_none_match = first$etag)
again$not_modified   # TRUE -- nothing was charged
```

## Errors

Problem+JSON responses become classed conditions:

```{r}
tryCatch(
  close_block_summary(close, "000000000000000"),
  close_api_error = function(e) cat(e$status, e$slug, "\n")
)
```

## Spatial output

With `sf` installed, convert a reply to an `sf` object:

```{r}
pts <- close_as_sf(close_pois_search(close, lat = 41.823, lon = -71.412,
                                     radius_m = 1500))
iso <- close_as_sf(close_isochrone(close, block = "440070036001010", minutes = 15))
```

POIs become points and isochrones become polygons offline; block replies join
census-block boundaries (pass `block_geometry =` or `fetch = TRUE`).
