---
title: "Get started with closecity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with closecity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Examples run when a key is present (on the docs site), and are shown but not run
# otherwise (on CRAN, and locally without a key). The displayed client uses a
# placeholder; the real one is built here from an environment variable.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

`closecity` reads the Close API: travel times from every US census block to nearby
places, on foot, by bike, and by public transit. This vignette is a short tour.
The three tutorials go further.

## Words you will see

A few terms come up throughout:

- **Census block.** The smallest area the Census Bureau publishes. Each one has a
  15-digit id, its **GEOID**.
- **Destination type.** A category of place, such as grocery stores or libraries.
  Every type has a numeric id.
- **Mode.** How someone travels: walk, bike, or transit.
- **Isochrone.** The area reachable from a point within a time limit, as a polygon.
- **Catchment.** The reverse: every block that can reach a given place.

## Build a client

You make every request through a client object.

```r
library(closecity)
close <- close_client("ck_live_your_key")   # use your own key here
```

The catalog and lookup routes are free, so `close_client()` with no key also works
for those.

```{r}
close$modes()$data$modes
```

## Look things up instead of guessing

Two free calls save you from memorising codes. Read the numeric id for a category
from the catalog, and turn a city name into a GEOID and a centre point.

```{r}
types <- close$destination_types()$data$destination_types
labels <- sapply(types, `[[`, "label")
ids <- sapply(types, `[[`, "dest_type_id")
grocery_id <- ids[labels == "grocery_stores"]

providence <- close$places("Providence")$results[[1]]
providence$geoid
```

## Make a call and map it

Feature methods return an [sf](https://r-spatial.github.io/sf/) object, so you can
map the result straight away.

```{r}
groceries <- close$pois_search(lat = providence$lat, lon = providence$lon,
                               radius_m = 1500, type = grocery_id)
plot(st_geometry(groceries), pch = 19, col = "#202a5b")
```

## Turn spatial output off

Set the client's `spatial` flag to FALSE to work with the raw data. Block routes
join census-block boundaries for you, using the `tigris` package.

```{r}
close$spatial <- FALSE
summary <- close$block_summary("440070036001010", mode = "walk")
summary$results
```

## Handle errors

Failed requests raise a classed condition. Catch the base `close_api_error`, or a
specific one.

```{r}
tryCatch(
  close$block_summary("000000000000000"),
  close_api_error = function(e) message(sprintf("%s (%d)", e$slug, e$status))
)
```
