---
title: "Get started with closecity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with closecity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Examples run when a key is present (on the docs site), and are shown but not run
# otherwise (on CRAN, and locally without a key). The displayed client uses a
# placeholder; the real one is built here from an environment variable.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

`closecity` reads the Close API: travel times from every US census block to nearby
places, on foot, by bike, and by public transit. This vignette is a short tour.
The three tutorials go further.

## Key terms

A few terms come up throughout:

- **Census block.** The smallest area the Census Bureau publishes. Each one has a
  15-digit id, its **GEOID**.
- **Destination type.** A category of place, such as grocery stores or libraries.
  Every type has a numeric id.
- **Mode.** How someone travels: walk, bike, or transit.
- **Isochrone** or **catchment.** Two views of the same reachability: the area
  reachable from a point within a time limit (an isochrone), or every block that
  can reach a given place (a catchment).

## Travel times

Times to nearby places are **capped at 30 minutes** for each mode, and recorded in
**whole minutes**. A missing time means the place is not reachable within the cap,
not that it is zero. Isochrones are the exception: they are available for any
budget up to an hour.

## Build a client

You make every request through a client object.

```r
library(closecity)
close <- close_client("ck_live_your_key")   # use your own key here
```

The catalog and lookup routes are free, so `close_client()` with no key also works
for those.

```{r}
close$modes()
```

## Look things up instead of guessing

Two free calls save you from memorising codes. Both come back as data frames, so
you filter and index them the usual way: read the numeric id for a category from
the catalog, and turn a city name into a GEOID and a centre point.

```{r}
types <- close$destination_types()
supermarket_dest_type <- types[types$label == "grocery_stores", ]$dest_type_id

providence_ri <- close$places("Providence")[1, ]
providence_ri[, c("name", "state", "geoid")]
```

The catalog's `name` column is the readable label ("Grocery stores"); the
underscored `label` is the internal key you match on. A place lookup carries a
`state`, so you can tell Providence, RI from the one in Utah.

## Make a call and map it

Routes with geometry return an [sf](https://r-spatial.github.io/sf/) object.
`close_map()` draws it on an interactive basemap in one line — bright, hoverable
points here.

```{r}
supermarkets <- close$place_pois(providence_ri$geoid, type = supermarket_dest_type)
close_map(supermarkets, color = "#e8590c")
```

## Choose an output

Every route returns tabular data by default: an sf object where geometry applies,
a data frame otherwise. The `output` setting changes the shape — `"tabular"` never
downloads boundaries, and `"raw"` gives the underlying reply with its metering and
cursor fields. Set it on the client, or pass `output =` to one call.

The same block summary as a data frame (the default):

```{r}
close$block_summary("440070008001068", mode = "walk")
```

...and as the raw reply, whose `results` you can index yourself:

```{r}
raw <- close$block_summary("440070008001068", mode = "walk", output = "raw")
head(raw$results, 3)
```

## Handle errors

Failed requests raise a classed condition. Catch the base `close_api_error`, or a
specific one.

```{r}
tryCatch(
  close$block_summary("000000000000000"),
  close_api_error = function(e) message(sprintf("%s (%d)", e$slug, e$status))
)
```
