---
title: "Competitor walksheds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Competitor walksheds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

**Question.** For a given business in a given industry, which competitors also serve
its walkshed — the residential blocks that can walk to it — and how much do their
catchments overlap? We use **Providence, RI** and the coffee-shop trade
(destination type `31`).

## Pick the business and its industry

```{r}
library(closecity)
close <- close_client("ck_live_your_key_here")

CAFE <- 31
providence <- close_places(close, "Providence")$results[[1]]

cafes <- close_records(close_pois_search, close,
                       lat = providence$lat, lon = providence$lon,
                       radius_m = 1200, type = CAFE)
ours <- cafes[[1]]
competitors <- cafes[2:6]
ours$name
```

## Our walkshed

Every block that can reach our shop within a 10-minute walk:

```{r}
walkshed <- function(dest_id) {
  rows <- close_records(close_poi_catchment, close, dest_id,
                        mode = "walk", max_minutes = 10)
  unique(vapply(rows, function(r) r$geoid, character(1)))
}

ours_shed <- walkshed(ours$dest_id)
length(ours_shed)
```

## Which competitors contest it

Local set math, free once the catchments are pulled:

```{r}
for (c in competitors) {
  shed <- walkshed(c$dest_id)
  shared <- intersect(ours_shed, shed)
  cat(sprintf("%-30s shares %3d blocks (%.0f%% of ours)\n",
              c$name, length(shared), 100 * length(shared) / length(ours_shed)))
}
```

## Map the contested ground

```{r}
library(sf)
pts <- close_as_sf(close_pois_search(close, lat = providence$lat,
                                     lon = providence$lon, radius_m = 1200,
                                     type = CAFE))
shed_blocks <- close_as_sf(close_poi_catchment(close, ours$dest_id, mode = "walk",
                                               max_minutes = 10), fetch = TRUE)

plot(st_geometry(shed_blocks), col = "#eef0f7", border = "#c6cbe0")
plot(st_geometry(pts), add = TRUE, col = "#202a5b", pch = 19)
plot(st_geometry(pts[pts$dest_id == ours$dest_id, ]), add = TRUE,
     col = "#f36e21", pch = 19, cex = 1.6)
```

## Scaling to a county or state

The same recipe runs over a wider area — search a bounding box instead of a radius —
but catchment pulls are metered per block, so a county-wide sweep can run into
thousands of tokens. Keep `max_minutes` small and page in batches.

## Token cost

- `close_places` + one `close_pois_search`: free lookup + a few dozen tokens.
- Six 10-minute walk catchments (~50-150 blocks each): ~600 tokens.

Well inside a 5,000-token month for a single business and its near neighbours.
