---
title: "Competitor walksheds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Competitor walksheds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = nzchar(Sys.getenv("CLOSECITY_KEY"))
)
library(closecity)
library(sf)
close <- closecity::close_client(api_key = Sys.getenv("CLOSECITY_KEY"))
```

A coffee shop wants to know which competitors draw from the same neighbourhood it
does. Its **walkshed** is every residential block that can walk to it. This tutorial
finds the competitors — the cafes those same blocks can also walk to — and maps
them. The example city is Providence, Rhode Island.

*Running this tutorial uses about 2,000 tokens.*

## Set up

Read the cafe category id from the free catalog and find the city.

```r
library(closecity)
library(sf)
close <- closecity::close_client(api_key = "ck_live_your_key")   # use your own key here
```

```{r}
amenity_types <- close$destination_types()
ids <- setNames(amenity_types$dest_type_id, amenity_types$label)
cafe <- ids[["cafes"]]

city <- close$places(q = "Providence")[1, ]
city_boundary <- close$place_boundary(geoid = city$geoid)
```

## Find the shops and our walkshed

`$place_pois()` returns every cafe within the city's boundary — no radius to guess.
Pick one as the subject, then pull its walkshed: every block that can walk to it in
10 minutes.

```{r}
cafes <- close$place_pois(geoid = city$geoid, type = cafe)
ours <- cafes[1, ]
ours$name
cafes$is_ours <- cafes$dest_id == ours$dest_id

our_shed <- close$poi_catchment(dest_id = ours$dest_id, mode = "walk", max_minutes = 10)
walkshed <- sf::st_union(our_shed)
```

Draw the walkshed, with the cafes on top and our shop in orange.

```{r}
closecity::close_map(
  x = cafes,
  color = ifelse(cafes$is_ours, "#f36e21", "#202a5b"),
  label = "name",
  background = walkshed,
  background_color = "#74b9ff",
  boundary = city_boundary
)
```

## Who else those blocks can reach

A competitor is a cafe that the residents of *our* walkshed can also walk to. So take
the blocks in our walkshed, and for each of the nearest cafes, pull its walkshed:
if the two share blocks, those residents can reach it too. `our_shed$geoid` is just a
vector of block ids, so the overlap is a plain set intersection. (We check the
nearest cafes to keep the token cost down; the recipe scales to all of them.)

```{r}
our_geoids <- our_shed$geoid
others <- cafes[!cafes$is_ours, ]
dist_to_ours <- as.numeric(sf::st_distance(others, ours))
nearest <- others[order(dist_to_ours)[seq_len(min(15, nrow(others)))], ]

shared <- vapply(nearest$dest_id, function(id) {
  their_shed <- close$poi_catchment(dest_id = id, mode = "walk", max_minutes = 10)
  length(intersect(our_geoids, their_shed$geoid))
}, integer(1))

competing <- nearest[shared > 0, ]
for (i in order(shared, decreasing = TRUE)) {
  if (shared[i] == 0) next
  cat(sprintf(
    "%-28s %3d shared blocks (%.0f%% of ours)\n",
    nearest$name[i],
    shared[i],
    100 * shared[i] / length(our_geoids)
  ))
}

cafes$is_competitor <- cafes$dest_id %in% competing$dest_id
```

## Map the contested ground

Draw every cafe over our walkshed: our shop in orange, the competitors those blocks
can also reach in red, and the rest in navy. The red cafes are the ones fighting for
the same walk-in traffic.

```{r}
cafe_color <- ifelse(
  cafes$is_ours,
  "#f36e21",
  ifelse(cafes$is_competitor, "#e03131", "#202a5b")
)
closecity::close_map(
  x = cafes,
  color = cafe_color,
  label = "name",
  background = walkshed,
  background_color = "#74b9ff",
  boundary = city_boundary
)
```

The same recipe scales up: raise the cafe count you check, or compare whole cities by
pulling each one's cafes with `$place_pois()`.
