---
title: "Competitor walksheds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Competitor walksheds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

A coffee shop wants to know which competitors draw from the same neighbourhood it
does. Its **walkshed** is every residential block that can walk to it. This tutorial
finds nearby competitors and measures how much of that walkshed they share. The
example city is Providence, Rhode Island.

## Set up

Read the cafe category id from the free catalog and find the city centre.

```r
library(closecity)
library(sf)
close <- close_client("ck_live_your_key")   # use your own key here
```

```{r}
types <- close$destination_types()$data$destination_types
labels <- sapply(types, `[[`, "label")
ids <- sapply(types, `[[`, "dest_type_id")
cafe <- ids[labels == "cafes"]

city <- close$places("Providence")$results[[1]]
```

## Find the shops

Search for cafes near downtown. The result is an sf of points, so plot them and pick
one shop as the subject.

```{r}
cafes <- close$pois_search(lat = city$lat, lon = city$lon,
                           radius_m = 1200, type = cafe)
plot(st_geometry(cafes), pch = 19, col = "#202a5b")

ours <- cafes[1, ]
ours$name
```

## Our walkshed

Ask for every block that can reach our shop within a 10-minute walk. This comes back
as sf polygons, with the block boundaries downloaded once by `tigris`.

```{r}
our_shed <- close$poi_catchment(ours$dest_id, mode = "walk", max_minutes = 10)

plot(st_geometry(our_shed), col = "#eef0f7", border = "#c6cbe0")
plot(st_geometry(ours), add = TRUE, pch = 19, col = "#f36e21", cex = 1.6)
```

## Who else serves it

For each of the nearest competitors, pull their walkshed and count the blocks they
share with ours. `our_shed$geoid` is just a vector of block ids, so the overlap is a
plain set intersection.

```{r}
for (i in 2:6) {
  their_shed <- close$poi_catchment(cafes$dest_id[i], mode = "walk",
                                    max_minutes = 10)
  shared <- intersect(our_shed$geoid, their_shed$geoid)
  cat(sprintf("%-28s %3d shared blocks (%.0f%% of ours)\n",
              cafes$name[i], length(shared),
              100 * length(shared) / nrow(our_shed)))
}
```

## Map the contested ground

Draw our walkshed, then every cafe on top, with our shop highlighted. The clusters of
competitors inside the walkshed are the ones competing for the same walk-in traffic.

```{r}
plot(st_geometry(our_shed), col = "#eef0f7", border = "#c6cbe0")
plot(st_geometry(cafes), add = TRUE, pch = 19, col = "#202a5b")
plot(st_geometry(ours), add = TRUE, pch = 19, col = "#f36e21", cex = 1.6)
```

The same recipe works over a wider area: search a bounding box instead of a radius,
and loop over more shops.
