---
title: "Your first walkability map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Your first walkability map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

The quickest way to feel what Close gives you: read one block's travel times, map
the groceries around it, then draw how far you can walk from it. The example city
is Providence, Rhode Island.

*Running this tutorial uses about 90 tokens.*

## Set up

Build a client, then read what you need from the free catalog.

```r
library(closecity)
library(sf)
close <- close_client("ck_live_your_key")   # use your own key here
```

```{r}
types <- close$destination_types()
grocery <- types$dest_type_id[types$label == "grocery_stores"]

downtown <- close$places("Providence")[1, ]
```

## Read one block's travel times

Pick a block and ask how long it takes to walk to each kind of amenity. The result
is a small data frame, one row per category.

```{r}
summary <- close$block_summary("440070008001068", mode = "walk")
summary[, c("dest_type_id", "travel_time")]
```

## Map the groceries nearby

A radius search returns points, ready to map.

```{r}
groceries <- close$pois_search(lat = downtown$lat, lon = downtown$lon,
                               radius_m = 1200, type = grocery)
plot(st_geometry(groceries), pch = 19, col = "#058040")
```

## Draw how far you can walk

An isochrone is the headline map: the area you can reach on foot in 10, 20, and 30
minutes. The contours come back largest first, so drawing them paints the nearer
times on top.

```{r}
rings <- close$isochrone(block = "440070008001068", mode = "walk",
                         direction = "from", contours = c(10, 20, 30))
plot(rings["contour"], pal = hcl.colors(3, "YlGnBu", rev = TRUE),
     border = "white", main = "Walk time from downtown (minutes)")
```

## Walk versus transit

The same block, the same 30-minute budget, two modes side by side. It is the
clearest way to see what the bus buys you.

```{r, fig.width = 8, fig.height = 4}
walk <- close$isochrone(block = "440070008001068", mode = "walk",
                        direction = "from", minutes = 30)
transit <- close$isochrone(block = "440070008001068", mode = "transit",
                           direction = "from", minutes = 30)

par(mfrow = c(1, 2))
plot(st_geometry(walk), col = "#05804033", border = "#058040",
     main = "Walk, 30 min")
plot(st_geometry(transit), col = "#202a5b33", border = "#202a5b",
     main = "Transit, 30 min")
```

## Where to next

- **Looking for a home**: find blocks near several amenities at once, then narrow
  to a commute.
- **The amenity basket**: population-weighted walkability coverage across a whole
  city.
- **Competitor walksheds**: who else draws from your neighbourhood.
