---
title: "Your first walkability map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Your first walkability map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

The quickest way to feel what Close gives you: read one block's travel times, map
the supermarkets around it, then draw how far you can walk from it. The example
city is Providence, Rhode Island.

*Running this tutorial uses about 85 tokens.*

## Set up

Build a client, then read what you need from the free catalog.

```r
library(closecity)
close <- close_client("ck_live_your_key")   # use your own key here
```

```{r}
types <- close$destination_types()
supermarket_dest_type <- types[types$label == "grocery_stores", ]$dest_type_id

providence_ri <- close$places("Providence")[1, ]
```

## Read one block's travel times

Pick a block and ask how long it takes to walk to each kind of amenity. Join the
catalog's readable `name` and sort by time, so the nearest things are on top.

```{r}
walk_times <- close$block_summary("440070008001068", mode = "walk")
walk_times <- merge(walk_times, types[, c("dest_type_id", "name")],
                    by = "dest_type_id")
walk_times[order(walk_times$travel_time), c("name", "travel_time")]
```

## Map the supermarkets nearby

`close_map()` draws the result on an interactive basemap in one line — bright,
hoverable points here.

```{r}
supermarkets <- close$pois_search(lat = providence_ri$lat, lon = providence_ri$lon,
                                  radius_m = 1200, type = supermarket_dest_type)
close_map(supermarkets, color = "#e8590c")
```

## Draw how far you can walk

An isochrone is the headline map: the area you can reach on foot in 10, 20, and 30
minutes. Shade it by the `contour` minutes and the nearer times stand out.

```{r}
rings <- close$isochrone(block = "440070008001068", mode = "walk",
                         direction = "from", contours = c(10, 20, 30))
close_map(rings, fill = "contour")
```

## Walk versus transit

The same block and the same 30-minute budget, on foot and by bus — the clearest
way to see what transit buys you.

```{r}
walk <- close$isochrone(block = "440070008001068", mode = "walk",
                        direction = "from", minutes = 30)
transit <- close$isochrone(block = "440070008001068", mode = "transit",
                           direction = "from", minutes = 30)

close_map(walk, color = "#058040")
close_map(transit, color = "#202a5b")
```

## Where to next

- **Looking for a home**: find blocks near several amenities at once, then narrow
  to a commute.
- **The amenity basket**: population-weighted walkability coverage across a whole
  city.
- **Competitor walksheds**: who else draws from your neighbourhood.
