---
title: "Looking for a home"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Looking for a home}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = nzchar(Sys.getenv("CLOSECITY_KEY")))
library(closecity)
library(sf)
close <- close_client(Sys.getenv("CLOSECITY_KEY"))
```

Say you are moving to a new city and want to live somewhere walkable. Here you find
the blocks that are within a 10-minute walk of a grocery store, a 5-minute walk of a
restaurant, and a 20-minute walk of a frequent-transit stop. Then you narrow those
blocks to the overlap of two commutes. The example city is Somerville, Massachusetts.

## Set up

Build a client, then read the pieces you need from the free catalog instead of
memorising codes.

```r
library(closecity)
library(sf)
close <- close_client("ck_live_your_key")   # use your own key here
```

```{r}
# The catalog lists every category with its numeric id. Pull the ids you need.
types <- close$destination_types()$data$destination_types
labels <- sapply(types, `[[`, "label")
ids <- sapply(types, `[[`, "dest_type_id")

grocery <- ids[labels == "grocery_stores"]
restaurant <- ids[labels == "restaurants"]
transit <- ids[labels == "frequent_transit"]

# Turn the city name into a GEOID and a centre point.
city <- close$places("Somerville")$results[[1]]
```

## See what is around

Look at the raw ingredients first. Each search returns points, so plot the three
amenities as three layers.

```{r}
groceries <- close$pois_search(lat = city$lat, lon = city$lon,
                               radius_m = 3000, type = grocery)
restaurants <- close$pois_search(lat = city$lat, lon = city$lon,
                                 radius_m = 3000, type = restaurant)
stops <- close$pois_search(lat = city$lat, lon = city$lon,
                           radius_m = 3000, type = transit)

plot(st_geometry(restaurants), col = "#c6cbe0", pch = 19, cex = 0.6)
plot(st_geometry(groceries), col = "#058040", pch = 19, add = TRUE)
plot(st_geometry(stops), col = "#f36e21", pch = 17, add = TRUE)
```

## Find the blocks that qualify

Pull the per-block walk times for the whole city. The result is an sf table with one
row per (block, category); block boundaries come from `tigris`, downloaded once and
cached.

```{r}
blocks <- close$place_blocks(city$geoid, mode = "walk",
                             type = c(grocery, restaurant, transit))
```

Each amenity has its own rule, so pick the blocks that pass each one, then keep the
blocks that pass all three.

```{r}
near_grocery <- unique(blocks$geoid[blocks$dest_type_id == grocery &
                                      blocks$travel_time <= 10])
near_restaurant <- unique(blocks$geoid[blocks$dest_type_id == restaurant &
                                         blocks$travel_time <= 5])
near_transit <- unique(blocks$geoid[blocks$dest_type_id == transit &
                                      blocks$travel_time <= 20])

candidates <- intersect(intersect(near_grocery, near_restaurant), near_transit)
winners <- blocks[blocks$geoid %in% candidates, ]

plot(st_geometry(blocks), col = "#eef0f7", border = "white")
plot(st_geometry(winners), col = "#f36e21", border = NA, add = TRUE)
```

## Narrow to a shared commute

Suppose two of you work in different places. A transit isochrone from each workplace
shows how far each commute reaches. Both come back as polygons.

```{r}
work_a <- close$isochrone(lon = -71.0865, lat = 42.3625, mode = "transit",
                          direction = "from", minutes = 20)
work_b <- close$isochrone(lon = -71.0589, lat = 42.3555, mode = "transit",
                          direction = "from", minutes = 20)

plot(st_geometry(work_a), col = "#05804033", border = "#058040")
plot(st_geometry(work_b), col = "#f36e2133", border = "#f36e21", add = TRUE)
```

Keep the winning blocks that also sit inside both commutes. That short list is where
to look.

```{r}
both_commutes <- st_intersection(st_union(work_a), st_union(work_b))
shortlist <- winners[st_intersects(winners, both_commutes, sparse = FALSE)[, 1], ]

plot(st_geometry(winners), col = "#eef0f7", border = "white")
plot(st_geometry(shortlist), col = "#058040", border = NA, add = TRUE)
```
